---
title: "Healthcare Final Project"
author: Mengfei Lyu
output: html_document
---
## Data Processing

```{r eval=T, message=F, warning=FALSE}
library(RNHANES)
library(mice)
library(ggplot2)
```

Download list of files and variables
```{r eval=T, message=F, warning=FALSE}
files <- nhanes_data_files()
variables <- nhanes_variables()
```

Download the required data file
```{r eval=T, message=F, warning=FALSE}
examination <- nhanes_load_data("BMX_H", "2013-2014")
demographic <- nhanes_load_data("DEMO_H", "2013-2014")
lab1 <- nhanes_load_data("FASTQX_H", "2013-2014")
lab2 <- nhanes_load_data("TRIGLY_H", "2013-2014")
question <- nhanes_load_data("BPQ_H", "2013-2014")
diery <- nhanes_load_data("DR1TOT_H", "2013-2014")
```

```{r eval=T, message=F, warning=FALSE}
View(examination) 
View(demographic)
View(lab1)
View(lab2)
View(question)
View(diery)
```

Select target features and rename them
```{r eval=T, message=F, warning=FALSE}
#examination data
variables <- c("SEQN","BMXWT","BMXHT","BMXBMI")
examine <- examination[variables]
new.names <- c("id","weight","height","bmi")
names(examine) <- new.names

#demographic data
variables <- c("SEQN","RIAGENDR","RIDAGEYR")
demography <- demographic[variables]
new.names <- c("id","gender","age")
names(demography) <- new.names

#laboratory data
variables <- c("SEQN","PHQ020","PHQ030","PHQ060")
labora1 <- lab1[variables]
new.names <- c("id","sugar_cream","alcohol","supplement")
names(labora1) <- new.names

variables <- c("SEQN","LBDLDL")
labora2 <- lab2[variables]
new.names <- c("id","LDL_cholesterol")
names(labora2) <- new.names

#questionnaire data
variables <- c("SEQN","BPQ060","BPQ100D")
ques <- question[variables]
new.names <- c("id","cholesterol_check","cholesterol_medicine")
names(ques) <- new.names

#diery data
variables <- c("SEQN","WTDRD1","DRQSPREP","DRQSDIET","DR1TKCAL","DR1TPROT","DR1TCARB","DR1TSUGR","DR1TFIBE","DR1TTFAT","DR1TCHOL","DR1TATOC","DR1TRET","DR1TVARA","DR1TACAR","DR1TBCAR","DR1TCRYP","DR1TLYCO","DR1TLZ","DR1TVB1","DR1TVB2","DR1TNIAC","DR1TVB6","DR1TFOLA","DR1TFA","DR1TCHL","DR1TVB12","DR1TVC","DR1TVD","DR1TVK","DR1TCALC","DR1TPHOS","DR1TMAGN","DR1TIRON","DR1TZINC","DR1TCOPP","DR1TSODI","DR1TPOTA","DR1TSELE","DR1TCAFF","DR1TTHEO","DR1TALCO","DR1TMOIS","DR1_320Z")
diet <- diery[variables]
new.names <- c("id","dietary_weight","salt","on_diet","energy","protein","carbo","sugars","fiber","fat","cholesterol","ve","retinol","va","alpha-caro","beta-caro","beta-cryptoxanthin","lycopene","lutein_zeaxanthin","vb1","vb2","niacin","vb6","folate","folic_acid","choline","vb12","vc","vd","vk","calcium","phosphorus","magnesium","iron","zinc","copper","sodium","potassium","selenium","caffeine","theobromine","alcohol_gm","moisture","water")
names(diet) <- new.names
```

Join tables with id
```{r eval=T, message=F, warning=FALSE}
merge1 <- merge(examine, demography, by = "id")
merge2 <- merge(merge1, labora1, by = "id")
merge3 <- merge(merge2, labora2, by = "id")
merge4 <- merge(merge3, ques, by = "id")
merged <- merge(merge4, diet, by = "id")
```

Classify whether a respontant has high LDL-cholesterol. Measurement comes from  National Heart, Lung, and Blood Institute
```{r eval=T, message=F, warning=FALSE}
merged$high_cholesterol[merged$age <18 &  merged$LDL_cholesterol >= 130]  <- 1
merged$high_cholesterol[merged$age >=18 &  merged$LDL_cholesterol >= 160]  <- 1
merged$high_cholesterol[is.na(merged$high_cholesterol)]  <- 0
```

Check missing rate and filter out features with missing rate larger than 10%.
```{r eval=T, message=F, warning=FALSE}
total <- nrow(merged)
columns <- colnames(merged)
for (c in columns){
  num=sum(is.na(merged[c]))
  print(paste(c, round(num/total*100,2)))
}
```

According to the result, drop columns including cholesterol_check,cholesterol_medicine as well as id and LDL-cholesterol
```{r eval=T, message=F, warning=FALSE}
merged <- subset(merged, 
                 select=-c(id,LDL_cholesterol,cholesterol_check,cholesterol_medicine) )
```

Deal with missing data.
If dietary_weight is 0.0, that means the respondent didn't take the diery survey, hence these records need to be excluded.
```{r eval=T, message=F, warning=FALSE}
merged <- subset(merged, dietary_weight >0)
```

For other missing data, conduct imputation to get full dataset
```{r eval=T, message=F, warning=FALSE}
surveydata <- mice(merged, m = 5, maxit=2, seed = 0)
cholesterol <- complete(surveydata,1)
```

Transfer categorical data to numeric, create one-hot encoding feature to represent categorical features
```{r eval=T, message=F, warning=FALSE}
# convert all categorical features to factor
cholesterol <- transform(cholesterol, 
            gender = as.factor(gender),
            sugar_cream = as.factor(sugar_cream),
            alcohol = as.factor(alcohol),
            supplement = as.factor(supplement),
            salt = as.factor(salt),
            on_diet = as.factor(on_diet))


categorical <- c("gender","sugar_cream","alcohol","supplement","salt","on_diet")
ohedata <- cholesterol[categorical]

ohedata <- model.matrix(~ .+0, data=ohedata,contrasts.arg = lapply(ohedata, contrasts, contrasts=FALSE))

processed <- data.frame(ohedata,cholesterol)
# drop original categorical columns
processed <- subset(processed, select= -c(gender,sugar_cream,alcohol,supplement,salt,on_diet))

#rename one-hot encoding features
newnames <- c("male","female","add_sugar_cream","no_sugar_cream","alcohol","no_alcohol","supplement","no_supplement","salt_never","salt_rare","salt_occasional","salt_often","salt_unknown","on_diet","no_diet","diet_unknown")
names(processed)[1:16] <- newnames
```

Data Exploration
Visualization 1-High cholesterol ratio in sample
```{r eval=T, message=F, warning=FALSE}
ggplot(processed,  aes(x = high_cholesterol) ) +geom_bar()
```
From the plot, we can see the ratio of positive data points is much lower than the negative datapoints. This indicates that true positive rate is a more proper model evaluation measurement than accuracy because even the model classfies all datapoints to negative, it can still achieve a high accuracy.
  
Visualization 2-Gender distribution in sample
```{r eval=T, message=F, warning=FALSE}
ggplot(cholesterol,  aes(x = gender) ) +geom_bar()
```
The distribution of gender within sample population is approximately balanced. Hence the analysis can provide a reasonable insight to both gender of population.

Visualization 3-Age distribution in sample
```{r eval=T, message=F, warning=FALSE}
ggplot(processed, aes(age)) +geom_bar()
```
The distribution of age within sample population has a higher ratio of people older than 80. Hence the analysis result will provide a more profound insight to elder people within the whole population.

Rewrite processed data to txt file for machine learning model construction.
```{r eval=T, message=F, warning=FALSE}
write.table(processed, "mydata.txt", sep="\t",row.names = F)
```

Get class weight
```{r eval=T, message=F, warning=FALSE}
processed <- transform(processed, 
            high_cholesterol = as.factor(high_cholesterol))
```